Legends for supplementary material

### 2019-01-24 Manuscript/Figures/figure_legends.Rmd
updated legends

**Figure 1: Model Construction.** The linear genome was binned according to the resolution of the Hi-C data. TAD boundaries were then flanked by 1 resolution-unit on both sides of the boundary point. [??? The following is unclear. Show Y on the figure. Are you talking about overlap only?] The response vector Y used for classification was determined by whether or not a genomic bin overlapped with a flanked region. The positional coordinates of each functional genomic element, obtained from ENCODE, were then used to define the feature space of the models.

**Figure 2. The three predictor types considered when quantifying the association between TAD boundaries and genomic annotations.** The _overlap count_ (OC) predictors consider the total number of genomic features that overlapped with each genomic bin. The _overlap percent_ (OP) predictors were calculated by dividing the sum of all feature widths within each specific bin by the total bin width. The _distance_ predictors were calculated by measuring the distance (in bases) from the center of each genomic bin to the center of the nearest genomic feature.

**Figure 3: Ensemble Framework/Model Building Pipeline.** A diagram of the model building pipeline given an a combination of inputs from the ensemble framework. The data was split into a 7:3 training set to testing set ratio, the variable reduction technique of choice was used to minimize the number of uninformative predictors, and a random forest classification algorithm was performed. Each model was then validated on the same testing set.

**Figure 4: Techniques for addressing class imbalance.** Barplots illustrating the class imbalance problem featured across each of the four different resolutions that were analyzed. Minority classes represent the number of genomic bins that contain a TAD boundary, while the majority classes represent the number of genomic bins that do not contain a TAD boundary.

# Figure 5: 5kb Model Performances

Comparing model performances for TAD boundary data at 5kb resolution across each re-sampling technique. Each square plot represents a random forest classifier performed on 5kb resolution binned boundary data using a particular predictor type, variable reduction algorithm, and re-sampling technique evaluated across 4 different performance metrics.

# Figure 6: 25kb Model Performances

Comparing model performances for TAD boundary data at 25kb resolution across each re-sampling technique. Each square plot represents a random forest classifier performed on 25kb resolution binned boundary data using a particular predictor type, variable reduction algorithm, and re-sampling technique evaluated across 4 different performance metrics.

# Figure 7: 50kb Model Performances

Comparing model performances for TAD boundary data at 50kb resolution across each re-sampling technique. Each square plot represents a random forest classifier performed on 50kb resolution binned boundary data using a particular predictor type, variable reduction algorithm, and re-sampling technique evaluated across 4 different performance metrics.

# Figure 8: 100kb Model Performances

Comparing model performances for TAD boundary data at 100kb resolution across each re-sampling technique. Each square plot represents a random forest classifier performed on 10
+0kb resolution binned boundary data using a particular predictor type, variable reduction algorithm, and re-sampling technique evaluated across 4 different performance metrics.

**Figure 9: Comparing Performances** Comparing performances of random forest classification algorithms, across TAD boundary data at each resolution for models using distance-type predictors, with elastic-net regularization and random under-sampling. Performances were aggregated by taking the average of each metric across 50 iterations of random under-sampling. [??? Poor legend]

**Figure 10: Comparing Variable Importance.** Comparing variable importance for the top 15 most predictive functional genomic elements for each of the 4 different resolutions. The x-axis represents the standardized difference between the out-of-bag prediction accuracy after permutting each predictor variable, averaged across all trees. The greater the mean standardized difference, the more importance the predictor is to the model. [??? What are A/B/C/D panels?]

# Figure 11: Model Performances for all Resolutions

Comparing model performances for TAD boundary data for all 4 resolutions. The four regions represent (from left to right; top to bottom) data at 5kb, 25kb, 50kb, and 100kb resolutions. Within each region, the rows represent the set of re-sampling techniques, while the columns represent the set of performance metrics that the models were evaluated on. Each plot compares 2 sets of models; one using LASSO regularization and one using Elastic-Net regularization. Within each set, each bar represents the performance of a model with a specific predictor type; either overlap counts (OC) in red, overlap percent (OP) in green, or distance in blue.

>>>>

**Figure 1: Model Construction.** Cell line specific Hi-C matices were obtained from [@rao20143d].  The linear genome was then binned according to the intervals of the square-symmetric contact matrices at given resolutions (10 kb, 25 kb, 50 kb, and 100 kb). The coordinates of subsequent topologically associating domains were identified using the Arrowhead TAD-calling algorithm [@durand2016juicebox]. Each TAD coordinate was then flanked by a unit-specific amount (in basepairs) respective to the resolution of the corresponding contact matrix. The response vector Y used for classification was determined by whether a genomic bin overlapped with a flanked boundary region (Y=1) or not (Y=0). The spatial relationship between positional coordinates of various functional genomic elements, obtained from ENCODE, and the genomic bins were then used to define the feature space of the models.

**Figure 2: Ensemble Framework/Model Building Pipeline.** A diagram of the model building pipeline given a combination of inputs from the ensemble framework. For a particular cell line, contact matrix resolution, and set of features, the data was split into a 7:3 training:testing set ratio.  For the training set, recursive feature elimination was used to reduce the feature space and minimize the influence of uninformative predictors. A particular resampling technique was applied to the reduced training set to create balanced classes. A random forest classification algorithm was then performed and validated using the testing set.

**Figure 3. The three predictor types considered when quantifying the association between TAD boundaries and genomic annotations.** The _overlap count_ (OC) predictors consider the total number of genomic features that overlapped with each genomic bin. The _overlap percent_ (OP) predictors were calculated by dividing the sum of all feature widths within each specific bin by the total bin width. The _distance_ predictors were calculated by measuring the distance (in bases) from the center of each genomic bin to the center of the nearest genomic feature.

**Figure 4. Precision-Recall curves for unbalanced 10 kb data.** Precision-recall curves were used to initially compare performance of random forest models with different predictor types on unbalanced data at 10 kb resolution. GM12878 (solid line) and K562 (dashed line) cell lines were compared with predictor type-specific models which included overlap counts (blue), overlap percents (purple), log distance (green), and all three combined (red). For both cell lines, random forest models with log distance-type predictors were found to have higher AUPRC values.

**Figure 5. Model Performances on 10kb data using distance-type predictors for different number of features** Using RFE, performances of a series of random forest models with different numbers of the most predictive features for each cell line was compared using AUPRC. For convenience and computational efficiency, only specific subsets of features were considered (increasing powers of 2). For the 10 kb data, near optimal performance was achieved for both the GM12878 (red) and K562 cell lines (blue).

**Figure 6. Variable importance heatmap for 10 kb data across cell lines.** Predictive importance of the top 32 genomic features across cell lines for 10 kb data. Importance was discretized into quantiles, with features that were not congruent between cell lines filled in grey.

### 2019-01-24 Manuscript/Supplementary/supplementary_legends.Rmd
updated legends

@@

## Table 2: Class Distributions for Other Resolutions

Summary of the class distributions for each of the other resolution domain data sets (25 kb, 50 kb, and 100 kb), across each re-sampling technique. As with the 5 kb data, all re-sampling techniques yielded completely balanced classes. That is, there was a 1:1 relationship between the majority and minority classes after sampling took place.

## Table 3: Model Performances When Not Accounting for Class Imbalance

Results of various metrics used to assess predictive performances for models that were built on fully imbalanced datasets.

## Figure 1: Performances of Models with Both SMOTE-Balanced and Randomly Over-sampled Data.

Barplots measuring the predictive performance of models built on datasets using 50 iterations of both the SMOTE algorithm (A) and random over-sampling (B).  The performance metrics considered were accuracy, AUC, F1-Score, and MCC. Performances were compared across 4 resolutions (10 kb, 25kb, 50kb, and 100 kb), for each predictor type (OC, OP, and Distances).

## Table 4:

Results of various metrics used to assess predictive performances for models that were built on random under-sampled datasets.

## Table 5:

Results of various metrics used to assess predictive performances for models that were built on both SMOTE-balanced and random over-sampled datasets.

>>>>

**Figure 1. Distance Densities across cell lines and resolutions**
Density curves of distances (in millions of base pairs) from genomic bins to flanked TAD boundaries for GM12878 (left column) and K562 (right column) cell lines. The untransformed distances (red) are highly right skewed. After performing a $log_{2}$ tranformation (blue), the distances exhibit more normalized densities. The normalized distances were exhibited across all data resolution (rows).

**Figure 2. Precision-Recall curves for unbalanced 25 kb, 50kb, and 100kb data.** Precision-recall curves were used to initially compare performance of random forest models with different predictor types on unbalanced data at 25 kb, 50kb, and 100kb resolution. GM12878 (solid line) and K562 (dashed line) cell lines were compared with predictor type-specific models which included overlap counts (blue), overlap percents (purple), log distance (green), and all three combined (red). For both cell lines, and among all resolutions, random forest models with log distance-type predictors were found to have higher AUPRC values.

**Figure 3. Model Performances on 25 kb, 50kb, and 100kb data using distance-type predictors for different number of features** Using RFE, performances of a series of random forest models with different numbers of the most predictive features for each cell line was compared using AUPRC. For convenience and computational efficiency, only specific subsets of features were considered (increasing powers of 2). For each resolution, near optimal performance was achieved for both the GM12878 (red) and K562 cell lines (blue).

**Figure 4. Variable importance heatmap for 25 kb, 50kb, and 100kb data across cell lines.** Predictive importance of the top 32 genomic features across cell lines for 25 kb, 50kb, and 100kb resolutions. Importance was discretized into quantiles, with features that were not congruent between cell lines filled in grey.

**Figure 5: Imbalance ratios across resolutions and cell lines.** Imbalance ratios, defined as the ratio of minority class samples to majority class samples, were plotted across resolutions for cell lines GM12878 (red) and K562 (blue). The mean imbalance ratios across resolutions were found to be X1 (SD=Y2) and X2 (SD=Y2) for the GM12878 and K562 cell lines respectively.

@@
